Equalizer\\Programming and User Guide

### 2012-06-26 Developer/ProgrammingGuide/paper.tex
Collage update

\clearpage
\pagenumbering{roman}
\pagestyle{headings}
\tableofcontents
%\clearpage
\listoffigures \vfill
\section*{Revision History}
{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Revision for Equalizer 0.5\\
    1.4     & Nov 25, 2008 & Revision for Equalizer 0.6\\
    1.6     & Aug 07, 2009 & Revision for Equalizer 0.9\\
    1.8     & Mar 21, 2011 & Revision for Equalizer 1.0\\
    1.10    & Feb 17, 2012 & Revision for Equalizer 1.2\\
    1.11.1  & Apr 10, 2012 & Minor revision\\
    1.11.2  & Jun 25, 2012 & Added \sref{sScaling} and \sref{sROI}\\
    1.11.3  & Jun 26, 2012 & Update \aref{aFileFormat}\\
    \hline
  \end{tabularx}}
\clearpage

@@

The purpose of the \textsf{SystemWindow} processing method,
\textsf{processEvent}, is to perform window-system specific event
processing. For example, \textsf{AGLWindow::processEvent} calls
\textsf{aglUpdateContext} whenever a resize event is received. For
further, generic processing, the \textsf{Event} is passed on to
\textsf{Window::processEvent}. This \textsf{Event} no longer contains
the native event.

\begin{wrapfigure}{r}{.382\textwidth}
  \vspace{-2ex}\includegraphics[width=.382\textwidth]{images/eventFilter.pdf}
  {\caption{\label{fEventProcessing}Event Processing}}
\end{wrapfigure}
\textsf{Window::processEvent} is responsible for handling
the event locally and for translating it into a generic
\textsf{ConfigEvent}. The config events are send to the application
thread using \textsf{Config::sendEvent}.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/netNode.pdf}
  {\caption{\label{fNetNode}Communication between two Nodes}}
\end{wrapfigure}
The \textsf{co::LocalNode} is the specialization of the node for the given
process. It encapsulates the communication logic for connecting remote nodes, as
well as object registration and mapping. Local nodes are set up in the listening
state during initialization.

@@

@@

Node packets are directly dispatched using
\textsf{Dispatcher::dispatchCommand}. For object packets, the appropriate object
is found on the local node and \textsf{LocalNode::dispatchCommand} will dispatch
the packet.

@@

@@

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/slaveWrites.pdf}
  {\caption{\label{fSlaveCommit}Slave Commit Communication Sequence}}
\end{wrapfigure}
When using \textsf{commit} on slave object instances, multiple modifications on
the same database may happen simultaneously. To avoid excessive synchronization
in the network layer, Equalizer implements a lightweight approach which requires
the application to avoid or handle conflicting changes on the same object.

@@

The master instance can \textsf{sync} delta data received from slaves. When
using the version generated by the slave object commit, only the delta for this
version is applied. When using \textsf{VERSION\_NEXT} as the version parameter
to \textsf{sync}, one slave commit is unpacked. This call may block if no data
is pending. When using \textsf{VERSION\_HEAD} all pending versions are
unpacked. The data is synchronized in the order it has been received from the
slaves. If a specific order is needed, it is the application's responsibility to
serialize the slave commits accordingly.

@@

@@

The \textsf{co::Barrier} provides a networked barrier primitive. It is an
\textsf{co::Object} used by Collage for software swap barrier support, but it
can be used as a generic barrier in application code.

@@

\subsection{\label{sNetUsage}Usage in Equalizer}

\subsubsection{\label{sEqObject}co::Serializable}

Equalizer uses the \textsf{co::Object} data distribution mechanism for
implementing data distribution between the server, application and render
clients. Using dirty bits, \textsf{co::Serializable} implements one
usage pattern for the \textsf{co::Object}, allowing inheritance with data
distribution. It is easier to use, but imposes one typical way to implement data
distribution.

The \textsf{co::Serializable} data distribution is based on the concept
of dirty bits. Dirty bits form a 64-bit mask which marks which parts of the
object have to be distributed during the next commit.

\begin{description}
\item[Inherit from \textsf{co::Serializable}:] The base class will
  provide the dirty bit management and call serialize and deserialize
  appropriately. By overriding \textsf{get\-Chan\-ge\-Type}, the default
  versioning strategy might be changed.
\item[Define new dirty bits:] Define dirty bits for each data item by
  starting at \textsf{Serializable::DIRTY\_CUSTOM}, shifting this value
  consecutively by each new dirty bit.
\item[Implement serialize and deserialize:] For each object-specific
  dirty bit which is set, stream the corresponding data item to or from
  the provided stream. Call the parent method first in both
  functions. For application-specific objects, write a (de-)serialization
  function.
\item[Mark dirty data:] In each 'setter' method, call \textsf{setDirty}
  with the corresponding dirty bit.
\end{description}

The Equalizer entities use the common base class
\textsf{eq::fabric::Object}. This class manages common entity data, and is not
intended to be subclassed directly.  The inheritance chain \textsf{eq::View}
$\rightarrow$ \textsf{eq::fabric::View} $\rightarrow$
\textsf{eq::fabric::Object} $\rightarrow$ \textsf{eq::fabric::Se\-rial\-izable}
$\rightarrow$ \textsf{co::Object} serves as sample implementation for this
process.

The registration and mapping of \textsf{co::Serializable}s is done in
the same way as for \textsf{co::Object}s, which has been described in
\sref{sNetObject2}. Equalizer objects derived from
\textsf{co::Serializable}, e.g., \textsf{eq::View}, are registered,
mapped and synchronized by Equalizer.

>>>>

\clearpage
\pagenumbering{roman}
\pagestyle{headings}
\tableofcontents
%\clearpage
\listoffigures \vfill
\section*{Revision History}
{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Revision for Equalizer 0.5\\
    1.4     & Nov 25, 2008 & Revision for Equalizer 0.6\\
    1.6     & Aug 07, 2009 & Revision for Equalizer 0.9\\
    1.8     & Mar 21, 2011 & Revision for Equalizer 1.0\\
    1.10    & Feb 17, 2012 & Revision for Equalizer 1.2\\
    1.11.1  & Apr 10, 2012 & Minor revision\\
    1.11.2  & Jun 25, 2012 & Add \sref{sScaling} and \sref{sROI}\\
    1.11.3  & Jun 26, 2012 & Add \sref{sZeroconf}, update \sref{sNetObject2} and \aref{aFileFormat}\\
    \hline
  \end{tabularx}}
\clearpage

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/eventFilter.pdf}
  {\caption{\label{fEventProcessing}Event Processing}}
\end{wrapfigure}
The purpose of the \textsf{SystemWindow} processing method,
\textsf{processEvent}, is to perform window-system specific event
processing. For example, \textsf{AGLWindow::processEvent} calls
\textsf{aglUpdateContext} whenever a resize event is received. For
further, generic processing, the \textsf{Event} is passed on to
\textsf{Window::processEvent}. This \textsf{Event} no longer contains
the native event.

\textsf{Window::processEvent} is responsible for handling
the event locally and for translating it into a generic
\textsf{ConfigEvent}. The config events are send to the application
thread using \textsf{Config::sendEvent}.

@@

The \textsf{co::LocalNode} is the specialization of the node for the given
process. It encapsulates the communication logic for connecting remote nodes, as
well as object registration and mapping. Local nodes are set up in the listening
state during initialization.

@@

\subsubsection{\label{sZeroconf}Zeroconf Discovery}

Each \textsf{LocalNode} provides a \textsf{co::Zeroconf} communicator, which
allows node and resource discovery. This requires compiling using a Lunchbox
library supporting dns\_sd. The service "\_collage.\_tcp" is used to announce
the presence of a listening \textsf{LocalNode} using the zeroconf protocol to
the network, unless the local node has no listening connections. The node
identifier and all listening connection descriptions are announced using keys
starting with "co\_". Internally Collage uses this information to connect
unknown nodes by using the node identifier alone.

Applications may use the zeroconf communicator to add additional key-value pairs
to announce application-specific data, and to retrieve a snapshot of all
key-value pairs of all discovered nodes on the network.

\subsubsection{Communication between Nodes}

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/netNode.pdf}
  {\caption{\label{fNetNode}Communication between two Nodes}}
\end{wrapfigure}
Node packets are directly dispatched using
\textsf{Dispat\-cher::\-dis\-patchCommand}. For object packets, the appropriate
object is found on the local node and \textsf{LocalNode::dis\-patchCommand} will
dispatch the packet.

@@

Objects may have a maximum number of unapplied versions, which will cause the
commit on the master instance to block if any slave instance has reached the
maximum number of queued versions. By default, slave instances can queue an
unlimited amount of unapplied versions.

Objects use compression plugins to reduce the amount of data sent over the
network. By default, the plugin with the highest compression ratio and lossless
compression for \textsf{EQ\_COMPRESSOR\_DATATYPE\_BYTE} tokens is chosen. The
application may override \textsf{Object::chooseCompressor} to deactivate
compression by returning \textsf{EQ\_COMPRESSOR\_NONE} or to select
object-specific compressors.

@@

\subsubsection{\label{sSerializable}co::Serializable}

\textsf{co::Serializable} implements one typical usage pattern for data
distribution bsed on \textsf{co::Object}. The \textsf{co::Serializable} data
distribution is founded on the concept of dirty bits, allowing inheritance with
data distribution. Dirty bits form a 64-bit mask which marks which parts of the
object have to be distributed during the next commit. It is easier to use, but
imposes one typical way to implement data distribution.

\begin{description}
\item[Inherit from \textsf{co::Serializable}:] The base class will
  provide the dirty bit management and call serialize and deserialize
  appropriately. By overriding \textsf{get\-Chan\-ge\-Type}, the default
  versioning strategy might be changed.
\item[Define new dirty bits:] Define dirty bits for each data item by
  starting at \textsf{Serializable::DIRTY\_CUSTOM}, shifting this value
  consecutively for each new dirty bit.
\item[Implement serialize and deserialize:] For each object-specific
  dirty bit which is set, stream the corresponding data item to or from
  the provided stream. Call the parent method first in both
  functions. For application-specific objects, write a (de-)serialization
  function.
\item[Mark dirty data:] In each 'setter' method, call \textsf{setDirty}
  with the corresponding dirty bit.
\end{description}

The registration and mapping of \textsf{co::Serializable}s is done in the same
way as for \textsf{co::Object}s, which has been described in
\sref{sNetObject2}.

@@

When using \textsf{commit} on slave object instances, multiple modifications on
the same database may happen simultaneously. To avoid excessive synchronization
in the network layer, Equalizer implements a lightweight approach which requires
the application to avoid or handle conflicting changes on the same object.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/slaveWrites.pdf}
  {\caption{\label{fSlaveCommit}Slave Commit Communication Sequence}}
\end{wrapfigure}
The master instance can \textsf{sync} delta data received from slaves. When
using the version generated by the slave object commit, only the delta for this
version is applied. When using \textsf{VERSION\_NEXT} as the version parameter
to \textsf{sync}, one slave commit is unpacked. This call may block if no data
is pending. When using \textsf{VERSION\_HEAD} all pending versions are
unpacked. The data is synchronized in the order it has been received from the
slaves. If a specific order is needed, it is the application's responsibility to
serialize the slave commits accordingly.

@@

\subsubsection{\label{sNetUsage}Usage in Equalizer}

Equalizer uses the common base class \textsf{eq::fabric::Object}, which derives
from \textsf{co::Se\-ria\-lizable} for implementing data distribution between
the server, application and render clients. This class manages common entity
data, and is not intended to be subclassed directly.  The inheritance chain
\textsf{eq::View} $\rightarrow$ \textsf{eq::fabric::View} $\rightarrow$
\textsf{eq::fabric::Ob\-ject} $\rightarrow$ \textsf{co::Se\-rial\-izable}
$\rightarrow$ \textsf{co::Object} serves as sample implementation for this
process.

Equalizer objects derived from \textsf{co::Serializable}, e.g.,
\textsf{eq::View}, are registered, mapped and synchronized by Equalizer. To
execute the rendering tasks, the server send the task command packet to the
render clients, which dispatch and execute them in the received order.

@@

The \textsf{co::Barrier} provides a networked barrier primitive. It is an
\textsf{co::Object} used by Equalizer to implement software swap barriers, but
it can be used as a generic barrier in application code.

@@
